---
title: "R Notebook"
output: pdf_
---
```{r}
rm(list = ls())
```

```{r}
devtools::install_github("hodgettsp/cesR")
devtools::install_github("mccormackandrew/mapcan")
install.packages("labelled")
install.packages("tidyverse")
install.packages("haven")
install.packages("here")
install.packages("gtsummary")
install.packages("nnet")
install.packages("tidybayes")
install.packages("lme4")
install.packages("brms")
install.packages("rstanarm")
devtools::install_github("paleolimbot/electionca")

```
1) read in the poll;
2) model the poll;
3) read in the post-stratification data; and
4) apply the model to the post-stratification data.
```{r}
library(cesR)
library(dplyr)
library(haven)
library(labelled)
library(tidyverse)
library(here)
library(gtsummary)
library(nnet)
library(mapcan)
library(lme4)
library(brms)
library(tidybayes)
library(rstanarm)
```

```{r}
# Grab the datasets
raw_web_data <- 
  read_dta(here("2019 Canadian Election Study - Online Survey v1.0.dta"))
```

```{r}
raw_web_data <- 
  labelled::to_factor(raw_web_data)
head(raw_web_data)
```


Signicant factors to look out for: province, age (cps19_age), gender, education, race, cps19_vote_2015, (cps_language_68/69 maybe (think French vs English))

Others to consider, income, premier (this is just province lol), constituencynumber, 

```{r}
# To grab the necessary variables to build our model
reduced_data <- 
  raw_web_data %>%
  select(
    cps19_gender,
    cps19_province,
    cps19_age,
    cps19_education,
    cps19_votechoice,
    cps19_v_likely,       # Willingness to vote in 2019 election
    cps19_language_68,    # English
    cps19_language_69,    # French
    constituencynumber,
    cps19_income_cat,     # income category (i.e 50 000 - 99 999)
    cps19_income_number,   # income exact number 
    cps19_citizenship
  )
```

```{r}
# We want to group some of the variables into one.
reduced_data <-
  reduced_data %>%
  mutate(
    cps19_age_group = case_when(
      cps19_age >= 18 & cps19_age <= 34 ~ "18 to 34",
      cps19_age >= 35 & cps19_age <= 54 ~ "35 to 54",
      cps19_age >= 55 & cps19_age <= 64 ~ "55 to 64",
      cps19_age >= 65 ~ "65+",
    )
  )
# We also do not want to consider individuals who are not interested in voting
# reduced_data <-
#   reduced_data %>%
#   filter(reduced_data$cps19_v_likely
#                            !="Unlikely to vote" &
#                              reduced_data$cps19_v_likely
#                            !="Certain not to vote"&
#                              reduced_data$cps19_v_likely
#                            !="I am not eligible to vote")
# 

reduced_data <-
  reduced_data %>%
  filter(reduced_data$cps19_v_likely
                           !="I am not eligible to vote")
unique(reduced_data["cps19_citizenship"]) # We only want Canadian citizens as others can't vote.
# Luckily the previous operation removed Permanant Residents and other similar types.

# Now get rid of NA values in constituencynumber
reduced_data <-
  reduced_data %>% 
  drop_na(constituencynumber)
```

```{r}
# Fix up the labels for cps19_gender
reduced_data <-
  reduced_data %>%
    mutate(
    cps19_gender = case_when(
      cps19_gender == "A woman" ~ "Female",
      cps19_gender == "A man" ~ "Male",
      cps19_gender == "Other (e.g. Trans, non-binary, two-spirit, gender-queer)" ~ "Other")
  )
```

```{r}
# We also would like to get rid of NA values for people who have not indicated who they wanted
# to vote for. Also get rid of Don't know and Another Party (since other parties have no chance
# of winning a seat)
reduced_data <-
  reduced_data %>%
    filter(reduced_data$cps19_votechoice !="",
           reduced_data$cps19_votechoice !="Don't know/ Prefer not to answer",
           reduced_data$cps19_votechoice !="Another party (please specify)")


```

```{r}
# Fix the education tags.
reduced_data <- 
  reduced_data %>% 
  filter(reduced_data$cps19_education !="Don't know/ Prefer not to answer")

reduced_data <-
  reduced_data %>%
  mutate(
    cps19_education = case_when(
      cps19_education == "No schooling" ~ "High School or Lower",
      cps19_education == "Some elementary school" ~ "High School or Lower",
      cps19_education == "Completed elementary school" ~ "High School or Lower",
      cps19_education == "Some secondary/ high school" ~ "High School or Lower",
      cps19_education == "Completed secondary/ high school" ~ "High School or Lower",
      cps19_education == "Some university" ~ "High School or Lower",
      cps19_education == "Some technical, community college, CEGEP, College Classique" ~ "Technical, college, CEGEP or other non-university certificate",
      cps19_education == "Completed technical, community college, CEGEP, College Classique" ~ "Technical, college, CEGEP or other non-university certificate",
      cps19_education == "Bachelor's degree" ~ "University degree", 
      cps19_education == "Master's degree" ~ "University degree",
      cps19_education == "Professional degree or doctorate" ~ "University degree",
      TRUE ~ "What happened here?") 
  )
```

```{r}
# Rename the parties to be consistent
reduced_data <-
  reduced_data %>%
  mutate(
    cps19_votechoice = case_when(
      cps19_votechoice == "Conservative Party" ~ "CPC",
      cps19_votechoice == "Liberal Party" ~ "LPC",
      cps19_votechoice == "Green Party" ~ "GPC",
      cps19_votechoice == "ndp" ~ "NDP",
      cps19_votechoice == "People's Party" ~ "PPC")
  )
# Note that there are now 1348 NA's, which correspond to Bloc (since this dataset
# has the Bloc written with non UTF-8 characters which makes it hard to read)
reduced_data$cps19_votechoice[is.na(reduced_data$cps19_votechoice)] <- "BQ"
```

```{r}
# We can improve our model by applying 2015's election results for each riding.
riding_results <- mapcan::federal_election_results %>% 
  filter(election_year == "2015") %>% 
  select(riding_code, 
         party) %>% 
  rename(constituencynumber = riding_code) 
riding_results$constituencynumber = riding_results$constituencynumber

riding_results <-
  riding_results %>%
    mutate(
    party = case_when(
      party == "Conservative" ~ "CPC",
      party == "Liberal" ~ "LPC",
      party == "Green" ~ "GPC",
      party == "NDP" ~ "NDP",
      party == "Bloc" ~ "BQ")
  )
```


# End of Data Cleaning
```{r}
# Compile everything in here
final_data <-
  reduced_data %>%
    select(
    cps19_gender,
    cps19_province,
    cps19_age_group,
    cps19_education,
    cps19_votechoice,
    cps19_v_likely,       # Willingness to vote in 2019 election
    constituencynumber,
  )
```

```{r}
final_data <-
  inner_join(final_data, riding_results, by=c("constituencynumber"))
```

```{r}
# Alternative using 2019 election data instead
riding_results2019 <-
  read_csv(here::here("2019-election.csv"))

riding_results2019 <-
  riding_results2019 %>% 
rename(constituencynumber = riding_code) %>% 
  rename(party = forecast_winner)

final_data <-
  inner_join(final_data, riding_results2019, by=c("constituencynumber"))
```


```{r}
# Let us plot a simple bar chart comparing voters for each corresponding party according
# to the poll
final_data%>%
  ggplot(aes(y=cps19_votechoice, fill = cps19_votechoice))+
  geom_bar(stat="count", width = 0.6)+
  scale_x_continuous(name = "Vote count", 
                     )+
  ylab("Most Likely to Vote For")+
  scale_fill_manual(values=c("CPC"= "cadetblue3","LPC"= "hotpink3",
                             "GPC"= "navyblue", "NDP"= "gainsboro", 
                             "BQ"= "gray20", "PPC" = "springgreen3"), 
                    name = "Party")+
  theme_minimal()
```
Now we can create our models.
```{r}
final_data %>% 
  tbl_summary()
```

```{r}
# We'll model with a logit (logistic) model
 glm_model_LPC <-
   glm(cps19_votechoice=="LPC" ~ cps19_gender + cps19_age_group + cps19_education + constituencynumber + party, data = final_data, family = binomial(link="logit"))

glm_model_CPC <-
  glm(cps19_votechoice=="CPC" ~ cps19_gender + cps19_age_group + cps19_education
      + constituencynumber + party,
     data = final_data, family = "binomial")

glm_model_NDP <-
  glm(cps19_votechoice=="NDP" ~ cps19_gender + cps19_age_group + cps19_education
      + constituencynumber + party,
     data = final_data, family = "binomial")

glm_model_PPC <-
  glm(cps19_votechoice=="PPC" ~ cps19_gender + cps19_age_group + cps19_education
      + constituencynumber + party,
     data = final_data, family = "binomial")

glm_model_BQ <-
  glm(cps19_votechoice=="BQ" ~ cps19_gender + cps19_age_group + cps19_education
      + constituencynumber + party,
     data = final_data, family = "binomial")

glm_model_GPC <-
  glm(cps19_votechoice=="GPC" ~ cps19_gender + cps19_age_group + cps19_education
      + constituencynumber + party,
     data = final_data, family = "binomial")

# OTHER ALTERNATIVE DIFFERENT MODEL
#  glm_model_LPC <-
#    glm(cps19_votechoice=="LPC" ~ cps19_gender + cps19_age_group + cps19_education
#        + party, data = final_data, family = "binomial")
# 
# glm_model_CPC <-
#   glm(cps19_votechoice=="CPC" ~ cps19_gender + cps19_age_group + cps19_education
#       +  party,
#      data = final_data, family = "binomial")
# 
# glm_model_NDP <-
#   glm(cps19_votechoice=="NDP" ~ cps19_gender + cps19_age_group + cps19_education
#       + party,
#      data = final_data, family = "binomial")
# 
# glm_model_PPC <-
#   glm(cps19_votechoice=="PPC" ~ cps19_gender + cps19_age_group + cps19_education
#       + party,
#      data = final_data, family = "binomial")
# 
# glm_model_BQ <-
#   glm(cps19_votechoice=="BQ" ~ cps19_gender + cps19_age_group + cps19_education
#       + party,
#      data = final_data, family = "binomial")
# 
# glm_model_GPC <-
#   glm(cps19_votechoice=="GPC" ~ cps19_gender + cps19_age_group + cps19_education
#       + party,
#      data = final_data, family = "binomial")
```


```{r}
# Load the models. This was from brm models but were weaker predictors than the glm ones.
# LPC_model <- readRDS(here("output/brm_LPC.rds"))
# CPC_model <- readRDS(here("output/brm_CPC.rds"))
# NDP_model <- readRDS(here("output/brm_NDP.rds"))
```


```{r}
# Since cps19_votechoice is an unordered categorical response variable, we should use
# a multinomial logistic regression (another weaker model)
# final_data$cps19_votechoice <- relevel(factor(final_data$cps19_votechoice), ref = "LPC")
# multinomial_model <- multinom(cps19_votechoice ~ cps19_gender + cps19_age_group + 
#                                  cps19_education + party,
#                               data = final_data)
```
Now that we have our model, we should do some model diagonostics.

 

Now that we pretended to do model diagonostics, let grab the post-stratification data. We will
be using data from the 2016 Census.  

```{r}
census_data <-
  read_csv(here::here("census_data.csv"))
head(census_data)
```

# Let us apply post-stratification then.

```{r}
post_strat_census_data <- census_data

post_strat_census_data <- 
  post_strat_census_data %>% 
  count(GEO_NAME, cps19_gender, cps19_education, cps19_age_group, cps19_province, ALT_GEO_CODE) %>%
  group_by(GEO_NAME, cps19_gender, cps19_education, cps19_age_group, cps19_province, ALT_GEO_CODE)

post_strat_census_data <-
  post_strat_census_data %>%
  group_by(GEO_NAME) %>%
  mutate(prop = n/sum(n)) %>%
  ungroup()

```


```{r}
# Testing out the glm models instead (ALTERNATIVE)

post_strat_census_data <- rename(post_strat_census_data, constituencynumber = ALT_GEO_CODE)

# Add 2019 vote to data
post_strat_census_data <-
  inner_join(post_strat_census_data, riding_results2019, by=c("constituencynumber"))

pred_model <- data.frame(predict(glm_model_LPC, newdata = post_strat_census_data, type = "response"))
colnames(pred_model) = "LPC"
post_strat_census_data <- cbind(post_strat_census_data, pred_model)

pred_model <- data.frame(predict(glm_model_CPC, newdata = post_strat_census_data, type = "response"))
colnames(pred_model) = "CPC"
post_strat_census_data <- cbind(post_strat_census_data, pred_model)

pred_model <- data.frame(predict(glm_model_NDP, newdata = post_strat_census_data, type = "response"))
colnames(pred_model) = "NDP"
post_strat_census_data <- cbind(post_strat_census_data, pred_model)

pred_model <- data.frame(predict(glm_model_PPC, newdata = post_strat_census_data, type = "response"))
colnames(pred_model) = "PPC"
post_strat_census_data <- cbind(post_strat_census_data, pred_model)

pred_model <- data.frame(predict(glm_model_BQ, newdata = post_strat_census_data, type = "response"))
colnames(pred_model) = "BQ"
post_strat_census_data <- cbind(post_strat_census_data, pred_model)

pred_model <- data.frame(predict(glm_model_GPC, newdata = post_strat_census_data, type = "response"))
colnames(pred_model) = "GPC"
post_strat_census_data <- cbind(post_strat_census_data, pred_model)

riding_data <-
post_strat_census_data %>%
  mutate(LPC = LPC*prop,
         CPC = CPC*prop,
         NDP = NDP*prop,
         PPC = PPC*prop,
         BQ = BQ*prop,
         GPC = GPC*prop)  %>%
  group_by(constituencynumber) %>%
  summarise(LPC = sum(LPC),
            CPC = sum(CPC),
            NDP = sum(NDP),
            PPC = sum(PPC),
            BQ = sum(BQ),
            GPC = sum(GPC))

# ALTERNATIVE
# riding_data <-
# post_strat_census_data %>% 
#   group_by(constituencynumber) %>%
#   summarise(LPC = sum(LPC),
#             CPC = sum(CPC),
#             NDP = sum(NDP),
#             PPC = sum(PPC),
#             BQ = sum(BQ),
#             GPC = sum(GPC))

# Finally we take the max of the 6 parties to see the victor of each riding
party_pred <- colnames(riding_data)[max.col(riding_data[,-1], "first") + 1]
riding_data <- cbind(riding_data, party_pred)
# Get rid of unnecessary columns
riding_data <- riding_data[-c(2, 3, 4, 5, 6, 7)]

```

```{r}
# Rename the headers
riding_data <- rename(riding_data, riding_code = constituencynumber)
riding_data <- rename(riding_data, forecast_winner = party_pred)
```

```{r}
# We now have our results. Write to an excel file.
write.csv(riding_data, here::here("forecast-summary.csv"), row.names = FALSE)
```



```{r}
# Check the results against 2015 election.
sum(riding_data$forecast_winner == riding_results2019$party)
```
```{r}
fed_ridings <- mapcan(boundaries = ridings,
       type = standard)
colnames(riding_data)[colnames(riding_data) == "constituencynumber"] <- "riding_code"

fed_ridings <- inner_join(fed_ridings, riding_data, by = "riding_code")
colnames(fed_ridings)[colnames(fed_ridings) == "forecast_winner"] <- "party"

fed_ridings <- 
  fed_ridings %>% 
  select(long, lat, party, group, riding_code)


federal_map <- fed_ridings %>%
  ggplot(aes(x = long, y = lat, group = group, fill = party)) +
  geom_polygon() +
  coord_fixed() +
  theme_mapcan() +
  ggtitle("2021 Federal Electoral Predictions")

federal_map +
  scale_fill_manual(name = "Winning party",
                    values = c("steelblue", "blue", "green", "red", "orange")) 


```





